---
title: "WAcovariatesimulation"
author: "Sean Bellew"
date: "25/05/2020"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We first extract the 19 bioclimatic bariables from worldclim data using the getData function (raster), picking the location of Margaret River WA as our study area.

```{r,warning=FALSE}
library(raster)
library(usdm)
library(spatstat)
library(maptools)
library(Metrics)
# reproducibility
set.seed(2)
```

We can choose to eliminate some of these covariates that are highly collerated using vifstep

```{r}
bioclim <- raster::getData('worldclim', 
                           var = 'bio',
                           res = 0.5,
                           lon = 115,
                           lat = -34)
vifstep(bioclim,
        th = 5)
```

We use these 5 covariates for our model.

```{r}
e <- extent(116, 119, -34, -31)
bioclim <- crop(bioclim, e)
bioclim <- bioclim[[c(3, 8, 9, 13, 14)]]
bio3 <- subset(bioclim, 1, drop = TRUE)
bio8 <- subset(bioclim, 2, drop = TRUE)
bio9 <- subset(bioclim, 3, drop = TRUE)
bio13 <- subset(bioclim, 4, drop = TRUE)
bio14 <- subset(bioclim, 5, drop = TRUE)
```

Set up and scale our covariates

```{r}
s <- raster(ncol = ncol(bioclim), 
            nrow = nrow(bioclim), 
            xmn = xmin(bioclim), 
            xmx = xmax(bioclim), 
            ymn = ymin(bioclim), 
            ymx = ymax(bioclim))
r <- s
values(r) <- values(bio3)
names(r) <- 'bio3'
s <- addLayer(s, r)
values(r) <- values(bio8)
names(r) <- 'bio8'
s <- addLayer(s, r)
values(r) <- values(bio9)
names(r) <- 'bio9'
s <- addLayer(s, r)
values(r) <- values(bio13)
names(r) <- 'bio13'
s <- addLayer(s, r)
values(r) <- values(bio14)
names(r) <- 'bio14'
s <- addLayer(s, r)
s <- scale(s, center = TRUE, scale = TRUE)
```

We set up our grid next

```{r}
res <- 10 / 60
s.loc <- xyFromCell(s, 1:ncell(s))
windowarea <- (xmax(bioclim) - xmin(bioclim)) * (ymax(bioclim) - ymin(bioclim))
#set up background grid (background resolution = resolution(s)*gridfactor)
gridfactor <- res / mean(res(s))
print(gridfactor)
row <- ceiling(gridfactor * c((gridfactor + 1) / (2 * gridfactor) + 0:(nrow(s) / gridfactor - 1)))
col <- ceiling(gridfactor * c((gridfactor + 1) / (2 * gridfactor) + 0:(ncol(s) / gridfactor - 1)))
cellnum <- cellFromRowColCombine(s, row, col)
#cellnum <- 1:ncell(s)
sgrid.loc <- xyFromCell(s, cellnum)
sgrid.loctot <- coordinates(s)
sgrid.bio3 <- values(s)[, 'bio3'][cellnum]
sgrid.bio8 <- values(s)[, 'bio8'][cellnum]
sgrid.bio9 <- values(s)[, 'bio9'][cellnum]
sgrid.bio13 <- values(s)[, 'bio13'][cellnum]
sgrid.bio14 <- values(s)[, 'bio14'][cellnum]
X.back1 <- cbind(1,
                 sgrid.bio3,
                 sgrid.bio8,
                 sgrid.bio9,
                 sgrid.bio13,
                 sgrid.bio14)
X.back1tot <- cbind(1, values(s))
```

We want to select our parameter values so the max number of points over the window is greater than 1000. This was done rather ad hoc using the max value of the intensity function, factoring in the NA values also.

```{r}
X <- cbind(1,
           values(s)[, 'bio3'],
           values(s)[, 'bio8'],
           values(s)[, 'bio9'],
           values(s)[, 'bio13'],
           values(s)[, 'bio14'])
gridsizeratio = sum(!is.na(X)) / length(X)
parameter <- c(beta1 = 5, beta2 = 0.2, beta3 = 0.3, beta4 = -0.2, beta5 = 0.26, beta6 = -0.2)
values(s) <- exp(X %*% parameter)
maxlambda <- max(values(s), na.rm = TRUE)
print(maxlambda)
``` 

We then run $N$ simulations for 32400 background/quadrature points, and average over the fits. Spatstat has been used as a comparison, with the covariates normalised and formatted to use in ppm.

```{r,warning=FALSE,include=FALSE} 
#set initial values
N=10
totlrbadd=matrix(0,N,dim(X)[2])
totwlrfit=matrix(0,N,dim(X)[2])
totfitIPP=matrix(0,N,dim(X)[2])
totfitIPPBT=matrix(0,N,dim(X)[2])
totppmIPP=matrix(0,N,dim(X)[2])
maxllfitIPP=matrix(0,N)
maxllfitIPPBT=matrix(0,N)
maxllwlrfit=matrix(0,N)
maxlllrbadd=matrix(0,N)
maxllppmIPP=matrix(0,N)
tmlrbadd=0
tmwlrfit=0
tmfitIPP=0
tmppmIPP=0
tmfitIPPBT=0

#create covariate list for use in spatstat from the covariate raster

bio3<-scale(bio3,center=TRUE,scale=TRUE)
bio8<-scale(bio8,center=TRUE,scale=TRUE)
bio9<-scale(bio9,center=TRUE,scale=TRUE)
bio13<-scale(bio13,center=TRUE,scale=TRUE)
bio14<-scale(bio14,center=TRUE,scale=TRUE)
cov3<-as.im.RasterLayer(bio3)
cov8<-as.im.RasterLayer(bio8)
cov9<-as.im.RasterLayer(bio9)
cov13<-as.im.RasterLayer(bio13)
cov14<-as.im.RasterLayer(bio14)
covlist<-list(a=cov3,b=cov8,c=cov9,d=cov13,e=cov14)
 w <- as.owin(cov3)
 
 #negative Poisson likelihood function (gridded only)
 
 likelihoodIPP = function(param){
    lambda = exp(X.back1 %*% param)
    mu = lambda * windowarea/nrow(X.back1)
    logL.pp = sum(X.po %*% param,na.rm=TRUE) - sum(mu,na.rm=TRUE)
    
    (-1)*sum(logL.pp)
 }
 
  likelihoodIPPtot = function(param){
    lambda = exp(X.back1tot %*% param)
    mu = lambda * windowarea/nrow(X.back1tot)
    logL.pp = sum(X.po %*% param,na.rm=TRUE) - sum(mu,na.rm=TRUE)
    
    (-1)*sum(logL.pp)
 }
 
 #negative Poisson likelihood function (X.back specifies Berman Turner, with area.back giving the weights)

likelihoodIPPBT = function(param){
  lambda = exp(X.back %*% param)
  mu = lambda * area.back
  logL.pp = sum(X.po %*% param,na.rm=TRUE) - sum(mu,na.rm=TRUE)
  
  (-1)*sum(logL.pp)
}

likelihoodIPPBTtot = function(param){
  lambda = exp(X.backtot %*% param)
  mu = lambda * area.backtot
  logL.pp = sum(X.po %*% param,na.rm=TRUE) - sum(mu,na.rm=TRUE)
  
  (-1)*sum(logL.pp)
}
 
 
 #negative weighted logistic regression likelihood function
 weight=1000
likelihoodwlr <- function(param) {
    alpha<-windowarea/(weight*dim(X.back1)[1])
    loglikewlr <- -(sum(X.po%*% param,na.rm=TRUE)-sum(log(1+alpha*exp(X.po%*% param)),na.rm=TRUE)-sum(weight*log(1+alpha*exp(X.back1%*% param)),na.rm=TRUE))
}

likelihoodwlrtot <- function(param) {
    alpha<-windowarea/(weight*dim(X.back1tot)[1])
    loglikewlr <- -(sum(X.po%*% param,na.rm=TRUE)-sum(log(1+alpha*exp(X.po%*% param)),na.rm=TRUE)-sum(weight*log(1+alpha*exp(X.back1tot%*% param)),na.rm=TRUE))
}

#negative baddeley regression likelihood function

likelihoodblr <- function(param){
  dum<-nrow(X.back1tot)/windowarea
  loglikblr <- sum(log(exp(X.po%*%param)/(exp(X.po%*%param)+dum)))+sum(log(dum/(exp(X.back1tot%*%param)+dum)))
  loglikblr

}

for (i in 1:N){
  
  #sample presence points
  N.hpp <- rpois(1,maxlambda*windowarea)
  ind.hpp <- sample(1:ncell(s), size=N.hpp, replace=FALSE)
  loc.hpp <- s.loc[ind.hpp, ]
  lambda.hpp <- values(s)[ind.hpp]
 
  ind.ipp <- runif(N.hpp, 0,1) <= lambda.hpp/maxlambda
  N.ipp <- sum(ind.ipp,na.rm=TRUE)
  loc.ipp <- loc.hpp[ind.ipp, ]
  X.po <- X[ind.hpp[ind.ipp], ]
  X.poslope<-X.po[,2:length(parameter)]
  
#setting up Berman-Turner quadrature
  
  X.back <- rbind(X.po,X.back1)
  X.backtot <- rbind(X.po,X.back1tot)
  s.loc.ipp=as.ppp(loc.ipp,W=w)
sgrid.loc.ipp<-as.ppp(sgrid.loc,W=w)
datadummy<-quadscheme(s.loc.ipp,sgrid.loc.ipp,method="grid")
area.back <-w.quad(datadummy)

sgrid.loc.ipptot<-as.ppp(sgrid.loctot,W=w)
datadummytot<-quadscheme(s.loc.ipp,sgrid.loc.ipptot,method="grid")
area.backtot <-w.quad(datadummytot)
  
#fit Poisson likelihood
  
tm <- Sys.time()
fitIPP <- optim(par=array(0,length(parameter)),fn=likelihoodIPP,method="BFGS")
tmfitIPP=tmfitIPP+Sys.time()-tm
totfitIPP[i,]=fitIPP$par
maxllfitIPP[i,]<- -likelihoodIPPtot(totfitIPP[i,])-sum(log(1:N.ipp))

#fit Poisson likelihood BT
tm <- Sys.time()
fitIPPBT <- optim(par=array(0,length(parameter)),fn=likelihoodIPPBT,method="BFGS")
tmfitIPPBT=tmfitIPPBT+Sys.time()-tm
totfitIPPBT[i,]=fitIPPBT$par
maxllfitIPPBT[i,]<- -likelihoodIPPBTtot(totfitIPPBT[i,])-sum(log(1:N.ipp))

#put covariates into dataframe for use in glm

  tm <- Sys.time()
  X.polr <-cbind(X.poslope,array(1,dim(X.poslope)[1]))
  X.dpp <-cbind(X.back1[,2:dim(X.back1)[2]],array(0,dim(X.back1)[1]))
  mydata <- rbind(X.polr,X.dpp)
  offst <- array(log(dim(X.back1)[1]/windowarea),dim(mydata)[1])
  mydata <- as.data.frame(mydata)
  mydata$offst <- offst
  
#fit Baddeley Regression
  
  lrbadd <- glm(V6~sgrid.bio3+sgrid.bio8+sgrid.bio9+sgrid.bio13+sgrid.bio14+offset(-offst),data=mydata,family="binomial")
   tmlrbadd=tmlrbadd+Sys.time()-tm
  totlrbadd[i,]=coefficients(lrbadd)
  maxlllrbadd[i,]=likelihoodblr(totlrbadd[i,])

  #fit Weighted logistic regression
  
  tm <- Sys.time()
  wlrfit <- optim(par=array(0,length(parameter)),fn=likelihoodwlr,method="BFGS")
   tmwlrfit=tmwlrfit+Sys.time()-tm
  totwlrfit[i,]=wlrfit$par
  maxllwlrfit[i]<- -likelihoodwlrtot(totwlrfit[i,])-sum(log(1:N.ipp))
  
# fit using ppm
  
tm <- Sys.time() 
mydata2<-as.data.frame(rbind(X.po,X.back1))
mydata2<-mydata2[,2:dim(mydata2)[2]]
ppmIPP<-ppm(datadummy,trend=~sgrid.bio3+sgrid.bio8+sgrid.bio9+sgrid.bio13+sgrid.bio14,data=mydata2,method='mpl'
)
tmppmIPP=tmppmIPP+Sys.time()-tm
totppmIPP[i,]=coefficients(ppmIPP)
maxllppmIPP[i,]=logLik(ppmIPP)-sum(log(1:N.ipp))

}
plot(loc.ipp)
fitIPP1=colSums(totfitIPP)/N
wlrfit1=colSums(totwlrfit)/N
lrbadd1=colSums(totlrbadd)/N
ppmIPP1=colSums(totppmIPP)/N

```

```{r}
'IPP'
print(fitIPP1)
'weighted logistic regression'
print(wlrfit1)
'Baddeley logistic regression'
print(lrbadd1)
'ppm spatstat'
print(ppmIPP1)
```

This can be compared to the true parameter values

```{r}
print(parameter)
```

Given is a boxplot for the 1st slope value, for the covariate bio3.

```{r}
sdata <- summary(totfitIPP[,3])[c(1,2,4,5,6)]
boxfitIPP <- list(stats=matrix(c(sdata[[1]],quantile(totfitIPP[,3],0.05),sdata[[3]],quantile(totfitIPP[,3],0.95),sdata[[5]]),5,1),n=10)
sdata <- summary(totwlrfit[,3])[c(1,2,4,5,6)]
boxwlrfit <- list(stats=matrix(c(sdata[[1]],quantile(totwlrfit[,3],0.05),sdata[[3]],quantile(totwlrfit[,3],0.95),sdata[[5]]),5,1),n=10)
sdata <- summary(totlrbadd[,3])[c(1,2,4,5,6)]
boxlrbadd <- list(stats=matrix(c(sdata[[1]],quantile(totlrbadd[,3],0.05),sdata[[3]],quantile(totlrbadd[,3],0.95),sdata[[5]]),5,1),n=10)
sdata <- summary(totppmIPP[,3])[c(1,2,4,5,6)]
boxppmIPP <- list(stats=matrix(c(sdata[[1]],quantile(totppmIPP[,3],0.05),sdata[[3]],quantile(totppmIPP[,3],0.95),sdata[[5]]),5,1),n=10)
box<-list(stats=cbind(boxfitIPP$stats,boxwlrfit$stats,boxlrbadd$stats,boxppmIPP$stats),n=c(boxfitIPP$n,boxwlrfit$n,boxlrbadd$n,boxppmIPP$n))
bp<-bxp(box,log="y",show.names=FALSE)
bxp(box,log="y",show.names=FALSE)
axis(1,at=bp,labels=c("IPP","weighted log","Baddeley","ppm spatstat"))
abline(h=parameter[3],col="Red")

```

```{r}
summary_results <- list()
summary_results$IPP <- matrix(NA, length(summary(totfitIPP[, 1])), ncol(totfitIPP))
for(i in seq_len(ncol(totfitIPP))) {
  summary_results$IPP[, i] <- summary(totfitIPP[, i])
}

summary_results$IPPBT <- matrix(NA, length(summary(totfitIPPBT[, 1])), ncol(totfitIPPBT))
for(i in seq_len(ncol(totfitIPPBT))) {
  summary_results$IPPBT[, i] <- summary(totfitIPPBT[, i])
}

summary_results$wlr <- matrix(NA, length(summary(totwlrfit[, 1])), ncol(totwlrfit))
for(i in seq_len(ncol(totwlrfit))) {
  summary_results$wlr[, i] <- summary(totwlrfit[, i])
}

summary_results$lr <- matrix(NA, length(summary(totlrbadd[, 1])), ncol(totlrbadd))
for(i in seq_len(ncol(totlrbadd))) {
  summary_results$lr[, i] <- summary(totlrbadd[, i])
}

summary_results$ppm <- matrix(NA, length(summary(totppmIPP[, 1])), ncol(totppmIPP))
for(i in seq_len(ncol(totppmIPP))) {
  summary_results$ppm[, i] <- summary(totppmIPP[, i])
}

for(i in seq_len(length(summary_results))) {
  colnames(summary_results[[i]]) <- names(parameter)
}

for(i in seq_len(length(summary_results))) {
  rownames(summary_results[[i]]) <- names(summary(totppmIPP[, 1]))
}

# We can also calculate the rmse for each fit

rmse_results <- list()
rmse_results$IPP <- matrix(NA, N)
rmse_results$IPPBT <- matrix(NA, N)
rmse_results$lr <- matrix(NA, N)
rmse_results$lrbadd <- matrix(NA, N)
rmse_results$ppmipp <- matrix(NA, N)
for(i in 1:N) {
  rmse_results$IPP[i] <- rmse(parameter, totfitIPP[i, ])
  rmse_results$IPPBT[i] <- rmse(parameter, totfitIPPBT[i, ])
  rmse_results$lr[i] <- rmse(parameter, totwlrfit[i, ])
  rmse_results$lrbadd[i] <- rmse(parameter, totlrbadd[i, ])
  rmse_results$ppmipp[i] <- rmse(parameter, totppmIPP[i, ])
}
```

Given is also the total running time for each method for 32400 background/quadrature points

```{r}
'IPP'
tmfitIPP
'weighted logistic regression'
tmwlrfit
'Baddeley logistic regression'
tmlrbadd
'ppm spatstat'
tmppmIPP
```

The code below runs the analysis for multiple resolutions and plots the results.

```{r}
bioclim <- raster::getData('worldclim', 
                           var = 'bio',
                           res = 0.5,
                           lon = 115,
                           lat = -34)

e <- extent(116, 119, -34, -31)
bioclim <- crop(bioclim, e)
bioclim <- bioclim[[c(3, 8, 9, 13, 14)]]
bio3 <- subset(bioclim, 1, drop = TRUE)
bio8 <- subset(bioclim, 2, drop = TRUE)
bio9 <- subset(bioclim, 3, drop = TRUE)
bio13 <- subset(bioclim, 4, drop = TRUE)
bio14 <- subset(bioclim, 5, drop = TRUE)

s <- raster(ncol = ncol(bioclim), 
            nrow = nrow(bioclim), 
            xmn = xmin(bioclim), 
            xmx = xmax(bioclim), 
            ymn = ymin(bioclim), 
            ymx = ymax(bioclim))
r <- s
values(r) <- values(bio3)
names(r) <- 'bio3'
s <- addLayer(s, r)
values(r) <- values(bio8)
names(r) <- 'bio8'
s <- addLayer(s, r)
values(r) <- values(bio9)
names(r) <- 'bio9'
s <- addLayer(s, r)
values(r) <- values(bio13)
names(r) <- 'bio13'
s <- addLayer(s, r)
values(r) <- values(bio14)
names(r) <- 'bio14'
s <- addLayer(s, r)
s <- scale(s, center = TRUE, scale = TRUE)

#create covariate list for use in spatstat from the covariate raster
  
bio3 <- scale(bio3, center = TRUE, scale = TRUE)
bio8 <- scale(bio8,center=TRUE,scale=TRUE)
bio9 <- scale(bio9, center = TRUE, scale = TRUE)
bio13 <- scale(bio13, center = TRUE, scale = TRUE)
bio14 <- scale(bio14, center = TRUE, scale = TRUE)

covlist <- list(a = as.im.RasterLayer(bio3),
                b = as.im.RasterLayer(bio8),
                c = as.im.RasterLayer(bio9),
                d = as.im.RasterLayer(bio13),
                e = as.im.RasterLayer(bio14))
w <- as.owin(covlist[[1]])

X <- cbind(1,
           values(s)[, 'bio3'],
           values(s)[, 'bio8'],
           values(s)[, 'bio9'],
           values(s)[, 'bio13'],
           values(s)[, 'bio14'])
gridsizeratio = sum(!is.na(X)) / length(X)
parameter <- c(beta1 = 5, beta2 = 0.2, beta3 = 0.3, beta4 = -0.2, beta5 = 0.26, beta6 = -0.2)
maxlambda <- max(exp(X %*% parameter), na.rm = TRUE)

resolutions <- c(5, 10, 20) / 60 # resolutions to consider
N <- 10 # number of replications

windowarea <- (xmax(bioclim) - xmin(bioclim)) * (ymax(bioclim) - ymin(bioclim))

sgrid.loctot <- coordinates(s)
X.back1tot <- cbind(1, values(s))

result <- setNames(lapply(resolutions, function(res) {
  s.loc <- xyFromCell(s, 1:ncell(s))
  # set up background grid (background resolution = resolution(s)*gridfactor)
  gridfactor <- res / mean(res(s))
  print(gridfactor)
  row <- ceiling(gridfactor * c((gridfactor + 1) / (2 * gridfactor) + 0:(nrow(s) / gridfactor - 1)))
  col <- ceiling(gridfactor * c((gridfactor + 1) / (2 * gridfactor) + 0:(ncol(s) / gridfactor - 1)))
  cellnum <- cellFromRowColCombine(s, row, col)
  #cellnum <- 1:ncell(s)
  sgrid.loc <- xyFromCell(s, cellnum)
  sgrid.bio3 <- values(s)[, 'bio3'][cellnum]
  sgrid.bio8 <- values(s)[, 'bio8'][cellnum]
  sgrid.bio9 <- values(s)[, 'bio9'][cellnum]
  sgrid.bio13 <- values(s)[, 'bio13'][cellnum]
  sgrid.bio14 <- values(s)[, 'bio14'][cellnum]
  X.back1 <- cbind(1,
                   sgrid.bio3,
                   sgrid.bio8,
                   sgrid.bio9,
                   sgrid.bio13,
                   sgrid.bio14)
  
  # set initial values
  totlrbadd <- totwlrfit <- totfitIPP <- totfitIPPBT <- totppmIPP <- matrix(0, N, dim(X)[2])
  maxllfitIPP <- maxllfitIPPBT <- maxllwlrfit <- maxlllrbadd <- maxllppmIPP <- matrix(0, N)
  tmlrbadd <- tmwlrfit <- tmfitIPP <- tmppmIPP <- tmfitIPPBT <- 0
  
  # negative Poisson likelihood function (gridded only)
  
  likelihoodIPP <- function(param){
    lambda <- exp(X.back1 %*% param)
    mu <- lambda * windowarea / nrow(X.back1)
    logL.pp <- sum(X.po %*% param, na.rm = TRUE) - sum(mu, na.rm = TRUE)
    
    -sum(logL.pp)
  }
  
  likelihoodIPPtot <- function(param){
    lambda <- exp(X.back1tot %*% param)
    mu <- lambda * windowarea / nrow(X.back1tot)
    logL.pp <- sum(X.po %*% param, na.rm = TRUE) - sum(mu, na.rm = TRUE)
    
    -sum(logL.pp)
  }
  
  # negative Poisson likelihood function (X.back specifies Berman Turner, with area.back giving the weights)
  
  likelihoodIPPBT <- function(param){
    lambda <- exp(X.back %*% param)
    mu <- lambda * area.back
    logL.pp <- sum(X.po %*% param, na.rm = TRUE) - sum(mu, na.rm = TRUE)
    
    -sum(logL.pp)
  }
  
  likelihoodIPPBTtot <- function(param){
    lambda <- exp(X.backtot %*% param)
    mu <- lambda * area.backtot
    logL.pp <- sum(X.po %*% param, na.rm = TRUE) - sum(mu, na.rm = TRUE)
    
    -sum(logL.pp)
  }
  
  
  # negative weighted logistic regression likelihood function
  weight <- 1e3
  likelihoodwlr <- function(param) {
    alpha <- windowarea / (weight * dim(X.back1)[1])
    loglikewlr <- -(sum(X.po %*% param, na.rm = TRUE) - 
                      sum(log(1 + alpha * exp(X.po %*% param)), na.rm = TRUE) -
                      sum(weight * log(1 + alpha * exp(X.back1 %*% param)), na.rm = TRUE))
  }
  
  likelihoodwlrtot <- function(param) {
    alpha <- windowarea / (weight * dim(X.back1tot)[1])
    loglikewlr <- -(sum(X.po %*% param, na.rm = TRUE) -
                      sum(log(1 + alpha * exp(X.po %*% param)), na.rm = TRUE) - 
                      sum(weight * log(1 + alpha * exp(X.back1tot %*% param)), na.rm = TRUE))
  }
  
  # negative baddeley regression likelihood function
  
  likelihoodblr <- function(param){
    dum <- nrow(X.back1tot) / windowarea
    loglikblr <- sum(log(exp(X.po %*% param) / (exp(X.po %*% param) + dum)), na.rm = TRUE) + 
      sum(log(dum / (exp(X.back1tot %*% param) + dum)), na.rm = TRUE)
    loglikblr
    
  }
  
  for(i in 1:N) {
    # sample presence points
    N.hpp <- rpois(1, maxlambda * windowarea)
    ind.hpp <- sample(1:ncell(s), size = N.hpp, replace = FALSE)
    loc.hpp <- s.loc[ind.hpp, ]
    lambda.hpp <- exp(X %*% parameter)[ind.hpp]
    
    ind.ipp <- runif(N.hpp, 0, 1) <= lambda.hpp / maxlambda
    N.ipp <- sum(ind.ipp, na.rm = TRUE)
    loc.ipp <- loc.hpp[ind.ipp, ]
    X.po <- X[ind.hpp[ind.ipp], ]
    X.poslope <- X.po[, 2:length(parameter)]
    
    #setting up Berman-Turner quadrature
    
    X.back <- rbind(X.po, X.back1)
    X.backtot <- rbind(X.po, X.back1tot)
    s.loc.ipp <- as.ppp(loc.ipp, W = w)
    sgrid.loc.ipp <- as.ppp(sgrid.loc, W = w)
    datadummy <- quadscheme(s.loc.ipp, sgrid.loc.ipp, method = "grid")
    area.back <- w.quad(datadummy)
    
    sgrid.loc.ipptot <- as.ppp(sgrid.loctot, W = w)
    datadummytot <- quadscheme(s.loc.ipp, sgrid.loc.ipptot, method = "grid")
    area.backtot <-w.quad(datadummytot)
    
    # fit Poisson likelihood
    
    tm <- Sys.time()
    fitIPP <- optim(par = array(0, length(parameter)), fn = likelihoodIPP, method = "BFGS")
    tmfitIPP <- tmfitIPP + Sys.time() - tm
    totfitIPP[i, ] <- fitIPP$par
    maxllfitIPP[i, ]<- -likelihoodIPPtot(totfitIPP[i,]) - sum(log(1:N.ipp))
    
    # fit Poisson likelihood BT
    tm <- Sys.time()
    fitIPPBT <- optim(par = array(0, length(parameter)), fn = likelihoodIPPBT, method = "BFGS")
    tmfitIPPBT <- tmfitIPPBT + Sys.time() - tm
    totfitIPPBT[i, ] <- fitIPPBT$par
    maxllfitIPPBT[i,] <- -likelihoodIPPBTtot(totfitIPPBT[i,]) - sum(log(1:N.ipp))
    
    # put covariates into dataframe for use in glm
    
    tm <- Sys.time()
    X.polr <- cbind(X.poslope, array(1, dim(X.poslope)[1]))
    X.dpp <- cbind(X.back1[,2:dim(X.back1)[2]], array(0, dim(X.back1)[1]))
    mydata <- rbind(X.polr, X.dpp)
    offst <- array(log(dim(X.back1)[1] / windowarea), dim(mydata)[1])
    mydata <- as.data.frame(mydata)
    mydata$offst <- offst
    
    #fit Baddeley Regression
    
    lrbadd <- glm(V6 ~ sgrid.bio3 + sgrid.bio8 + sgrid.bio9 + sgrid.bio13 + sgrid.bio14 + offset(-offst), 
                  data = mydata,
                  family = "binomial")
    tmlrbadd <- tmlrbadd + Sys.time() - tm
    totlrbadd[i,] <- coefficients(lrbadd)
    maxlllrbadd[i, ] <- likelihoodblr(totlrbadd[i, ])
    
    # fit Weighted logistic regression
    
    tm <- Sys.time()
    wlrfit <- optim(par = array(0, length(parameter)), fn = likelihoodwlr, method = "BFGS")
    tmwlrfit <- tmwlrfit + Sys.time() - tm
    totwlrfit[i, ] <- wlrfit$par
    maxllwlrfit[i] <- -likelihoodwlrtot(totwlrfit[i, ]) - sum(log(1:N.ipp))
    
    # fit using ppm
    
    tm <- Sys.time() 
    mydata2 <- as.data.frame(rbind(X.po, X.back1))
    mydata2 <- mydata2[, 2:dim(mydata2)[2]]
    ppmIPP <- ppm(datadummy, 
                  trend = ~ sgrid.bio3 + sgrid.bio8 + sgrid.bio9 + sgrid.bio13 + sgrid.bio14, 
                  data = mydata2, 
                  method = 'mpl'
    )
    tmppmIPP <- tmppmIPP + Sys.time() - tm
    totppmIPP[i, ] <- coefficients(ppmIPP)
    maxllppmIPP[i, ] <- logLik(ppmIPP) - sum(log(1:N.ipp))
    
  }
  fitIPP1 <- colSums(totfitIPP) / N
  wlrfit1 <- colSums(totwlrfit) / N
  lrbadd1 <- colSums(totlrbadd) / N
  ppmIPP1 <- colSums(totppmIPP) / N
  
  summary_results <- list()
  summary_results$IPP <- matrix(NA, length(summary(totfitIPP[, 1])), ncol(totfitIPP))
  for(i in seq_len(ncol(totfitIPP))) {
    summary_results$IPP[, i] <- summary(totfitIPP[, i])
  }
  
  summary_results$IPPBT <- matrix(NA, length(summary(totfitIPPBT[, 1])), ncol(totfitIPPBT))
  for(i in seq_len(ncol(totfitIPPBT))) {
    summary_results$IPPBT[, i] <- summary(totfitIPPBT[, i])
  }
  
  summary_results$wlr <- matrix(NA, length(summary(totwlrfit[, 1])), ncol(totwlrfit))
  for(i in seq_len(ncol(totwlrfit))) {
    summary_results$wlr[, i] <- summary(totwlrfit[, i])
  }
  
  summary_results$lr <- matrix(NA, length(summary(totlrbadd[, 1])), ncol(totlrbadd))
  for(i in seq_len(ncol(totlrbadd))) {
    summary_results$lr[, i] <- summary(totlrbadd[, i])
  }
  
  summary_results$ppm <- matrix(NA, length(summary(totppmIPP[, 1])), ncol(totppmIPP))
  for(i in seq_len(ncol(totppmIPP))) {
    summary_results$ppm[, i] <- summary(totppmIPP[, i])
  }
  
  for(i in seq_len(length(summary_results))) {
    colnames(summary_results[[i]]) <- names(parameter)
  }
  
  for(i in seq_len(length(summary_results))) {
    rownames(summary_results[[i]]) <- names(summary(totppmIPP[, 1]))
  }
  
  # We can also calculate the rmse for each fit
  
  rmse_results <- list()
  rmse_results$IPP <- matrix(NA, N)
  rmse_results$IPPBT <- matrix(NA, N)
  rmse_results$lr <- matrix(NA, N)
  rmse_results$lrbadd <- matrix(NA, N)
  rmse_results$ppmipp <- matrix(NA, N)
  for(i in 1:N) {
    rmse_results$IPP[i] <- rmse(parameter, totfitIPP[i, ])
    rmse_results$IPPBT[i] <- rmse(parameter, totfitIPPBT[i, ])
    rmse_results$lr[i] <- rmse(parameter, totwlrfit[i, ])
    rmse_results$lrbadd[i] <- rmse(parameter, totlrbadd[i, ])
    rmse_results$ppmipp[i] <- rmse(parameter, totppmIPP[i, ])
  }
  
  list(summary = summary_results,
       rmse = rmse_results)
}), nm = resolutions)
```
